---
title: "Case Study 1 - Final Report"
author: "Ikenna Nwaogu & Sterling Beason"
date: "10/26/2019"
output: html_document
---

```{r setup, echo=FALSE, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# libraries
library(tidyverse)
library(dplyr)
library(plyr)
library(maps)
library(class)
library(caret)

# read in data
beers = read.csv('./data/Beers.csv', header = TRUE) # Import Beers data
brew = read.csv('./data/Breweries.csv',header = TRUE) # Import Breweries data
map = read.csv('./data/us-zip-code-latitude-and-longitude.csv',header = TRUE,sep=";")  # Import Zipcode USA Coordinates
```
### Acronyms
**Alcohol Content = ABV**

**International Bitterness Units = IBU**

**K Nearest Neighbors = KNN**

## Question 1: Breweries Per State
**Below are the frequency of breweries per state. Colorado has the most at 47.**
```{r, echo=FALSE}
brewState <- brew %>% group_by(State)%>% summarize(count = count(State))
knitr::kable(x = brewState$count, format="html")
```

## Question 2: Merge Beer and Breweries Data
### First and Last Six Observations
```{r, echo=FALSE}
#Observe that ID on beers - Brewery_id and and breweries - Brew_ID  - we rename by Brew_ID

mergeData = inner_join(brew,beers, by = c("Brew_ID" = "Brewery_id"))

mergeData <- dplyr::rename(mergeData, brewName = Name.x)
mergeData <- dplyr::rename(mergeData, beerName = Name.y)

# check merged data
head(mergeData) # first 6 observations
tail(mergeData) # last 6 observations
```

## Question 3: Missing Values in Each Column
**ABV and IBU both have missing values in the dataset. ABV and IBU have 62 and 1005 NA values respectively**
```{r, echo=FALSE}
colSums(is.na(mergeData))
```

## Question 4: Median ABV and IBU for Each State
**ABV has less variance than the IBU variable. Interestingly Maine has the highest median IBU**
```{r, echo=FALSE}
#we will create 2 df for ABV not NA and IBU not NA - we dont want 1 df for both because we loose a lot of data - if we considerABV,... only 62 deleted and IBU

# Plot ABV by state
mergeData.cleanABV <- mergeData %>% filter(!is.na(ABV))
MedianABV <- ddply(mergeData.cleanABV, .(State), summarise, medABV=median(ABV))
MedianABV %>% ggplot(aes(fill=State, y=medABV, x=State)) + geom_bar(stat="identity") + labs(title = 'Median ABV By State')

# Plot IBU by state
mergeData.cleanIBU <- mergeData %>%  filter(!is.na(IBU))
MedianIBU <- ddply(mergeData.cleanIBU, .(State), summarise, medIBU=median(IBU))
MedianIBU %>% ggplot(aes(fill=State, y=medIBU, x=State)) + geom_bar(stat="identity") + labs(title = 'Median IBU By State')
```

## Question 5: Highest ABV and IBU by State
**Colorado has the highest ABV and Oregon has the highest IBU.**
```{r, echo=FALSE}
# Find state with highest ABV
mergeData.cleanABV %>% filter(ABV == max(ABV)) %>% select(State,ABV)

# Find state with highest IBU
mergeData.cleanIBU %>% filter(IBU == max(IBU)) %>% select(State,IBU)
```

### Maps For Questions 1, 4, 5
```{r, echo=FALSE}
#map_data('usa')
#map()
usa = map_data('usa')

us_states <- map_data("state")

#map <- map %>% select(State, Latitude, Longitude) %>% group_by(State) %>% summarise(long = median(Longitude))
map <- ddply(map, .(State), summarise, Long=median(Longitude), Lat = median(Latitude))

coords <- map

# Add coordinate to merge data by join
mergeData.clean = mergeData.cleanABV  %>% filter(!is.na(IBU))
mergeData.clean$State <- str_trim(mergeData.clean$State)
mergeData.clean$State <- as.character(mergeData.clean$State)
coords$State <- as.character(coords$State)
newData <- inner_join(mergeData.clean,coords, by = c("State" = "State"))

brewState$count$x <- as.character(str_trim(brewState$count$x))

g = c(brewState$count$x) 
e = c(brewState$count$freq)
Q = data.frame(State = g, Count = e)

Q$State = as.character(Q$State)
Q1 = inner_join(newData,Q, by = c("State" = "State"))
Q1 = distinct(Q1,State,.keep_all = TRUE)
Q1range = Q1 %>% filter(Count == max(Count) | Count == min(Count)) #select only max an min count
Q1rangew = Q1 %>% filter(Count != max(Count) & Count != min(Count)) # filter out max and min
theme_set(theme_dark())

#### Map without Alaska and Hawaii
# LOWER 48 FREQ
Q1w = Q1 %>% filter(State != 'AK' & State != 'HI')
Q1wrange = Q1w %>% filter(Count == max(Count) | Count == min(Count)) # filter in only max and min
Q1wrangew = Q1w %>% filter(Count != max(Count) & Count != min(Count)) # filter out max and min
ggplot() + geom_polygon(data = us_states, aes(long,lat,group=group), fill = 'blue', color = 'black') + geom_text(data = Q1w,aes(Long,Lat, label = State),hjust = 0, nudge_x = 0, color = 'black', size = 2) + geom_text(data = Q1wrangew,aes(Long,Lat, label = Count),hjust = 0, nudge_x = 0.2, nudge_y = -0.3, color = 'white', size = 2.3) + geom_text(data = Q1wrange,aes(Long,Lat, label = Count),hjust = 0, nudge_x = 0.2, nudge_y = -0.3, color = 'red', size = 2.5) + labs(title = 'Brewery Frequency By State')

###### Using maps for the median data ########
MedianABV$State <- str_trim(as.character(MedianABV$State))
MedianIBU$State <- str_trim(as.character(MedianIBU$State))
Q4 = inner_join(newData,MedianABV, by = c("State" = "State"))
Q4 = inner_join(Q4,MedianIBU, by = c("State" = "State"))
Q4 = distinct(Q4,State,.keep_all = TRUE)
Q4ABVrange = Q4 %>% filter(medABV == max(medABV) | medABV == min(medABV)) # filter in max and min ABV
Q4IBUrange = Q4 %>% filter(medIBU == max(medIBU) | medIBU == min(medIBU)) # filter in max and min IBU

Q4ABVrangeWO = Q4 %>% filter(medABV != max(medABV) & medABV != min(medABV)) ## filter out min max ABV
Q4IBUrangeWO = Q4 %>% filter(medIBU != max(medIBU) & medIBU != min(medIBU)) # filter out min max IBU


# Medians
# Now lets take the outlier states to make the gragh look better.
Q4w = Q4 %>% filter(State != 'AK' & State != 'HI')
Q4wABVrange = Q4w %>% filter(medABV == max(medABV) | medABV == min(medABV)) #filter in min max
Q4wIBUrange = Q4w %>% filter(medIBU == max(medIBU) | medIBU == min(medIBU)) #Filter in min max IBU

Q4wABVrangeWO = Q4w %>% filter(medABV != max(medABV) & medABV != min(medABV)) #filter out min max
Q4wIBUrangeWO = Q4w %>% filter(medIBU != max(medIBU) & medIBU != min(medIBU)) #Filter out min max IBU

# LOW HIGH ABV MEDIANS
ggplot() + geom_polygon(data = us_states, aes(long,lat,group=group), fill = 'blue', color = 'black') + geom_text(data = Q4w,aes(Long,Lat, label = State),hjust = 0, nudge_x = 0, color = 'black', size = 2) + geom_text(data = Q4wABVrangeWO,aes(Long,Lat, label = medABV),hjust = 0, nudge_x = 0, nudge_y = -0.5, color = 'white', size = 2.3) + geom_text(data = Q4wABVrange,aes(Long,Lat, label = medABV),hjust = 0, nudge_x = 0, nudge_y = -0.5, color = 'red', size = 2.5) + labs(title = 'Median ABV By State (low/high highlighted red)')

# LOW HIGH IBU MEDIANS
ggplot() + geom_polygon(data = us_states, aes(long,lat,group=group), fill = 'blue', color = 'black') + geom_text(data = Q4w,aes(Long,Lat, label = State),hjust = 0, nudge_x = 0, color = 'black', size = 2) + geom_text(data = Q4wIBUrangeWO,aes(Long,Lat, label = medIBU),hjust = 0, nudge_x = 0.1, nudge_y = -0.5, color = 'white', size = 2.3) + geom_text(data = Q4wIBUrange,aes(Long,Lat, label = medIBU),hjust = 0, nudge_x = 0, nudge_y = -0.5, color = 'red', size = 2.5) + labs(title = 'Median IBU By State (low/high highlighted red)')

```

## Question 6: ABV Summary Statistics 
**The distribution of ABV across beers is slightly right-skewed.**

**Most beers’ ABV are within 2.5% and 10%.**

**The median ABV is 5.6% and the mean is 5.98%.**

**The max ABV is 12.8%.**
```{r, echo=FALSE, warning=FALSE, message=FALSE}
# ABV summary statistics
summary(mergeData.cleanABV$ABV)

# Histogram of ABV distribution
mergeData.cleanABV %>% select(ABV) %>% ggplot(mapping=aes(x = ABV)) + geom_histogram()
```

## Question 7: ABV vs IBU Scatterplot
**There is a positive relationship between alcohol content and the bitterness of a beer generally.**

**You can see the higher the ABV, the higher the IBU.**
```{r, echo=FALSE, warning=FALSE}
# Filter out observations with NA value in either ABV or IBU
# Plot points with linear relationship line drawn
beers %>% filter(!is.na(IBU)) %>% select(ABV, IBU) %>% ggplot(mapping = aes(ABV, IBU)) + geom_point() + stat_summary(fun.data=mean_cl_normal) + geom_smooth(method='lm')
```

# Question 8: Style Predictions from ABV and IBU in KNN
**We selected k = 6 after cross-validation using 100 iterations of train/test data split and 1-30 k values. With k = 6, accuracy = 81.2%, sensitivity = 87.8%, and specificity = 71.3%.**
```{r, echo=FALSE}

brewState <- brew %>% group_by(State)%>% summarize(count = count(State))
brew %>% filter(State == 'SD')

#2. Merge beer data with the breweries data. Print the first 6 observations and the last six observations to check the merged file.  (RMD only, this does not need to be included in the presentation or the deck.)



#Observe that ID on beers - Brewery_id and and breweries - Brew_ID  - we rename by Brew_ID

mergeData = inner_join(brew,beers, by = c("Brew_ID" = "Brewery_id"))
names(mergeData)

mergeData <- dplyr::rename(mergeData, brewName = Name.x)
mergeData <- dplyr::rename(mergeData, beerName = Name.y)

mergeData.clean <- mergeData %>% filter(!is.na(ABV) &!is.na(IBU))
view(mergeData.clean)

typeAles <- cbind(mergeData.clean, type='Ales', stringsAsFactors=F) %>% filter(grepl('\\bale\\b', Style, ignore.case=T) | grepl('\\bale\\b', beerName, ignore.case=T))

typeIPA <- cbind(mergeData.clean, type='IPA', stringsAsFactors=F) %>% filter(grepl('\\bIPA\\b', Style, ignore.case=T) | grepl('\\bIPA\\b', beerName, ignore.case=T))

IPA.Ales <- union(typeAles, typeIPA)
IPA.Ales$type <- as.factor(IPA.Ales$type)



#KNN model
set.seed(100)
splitPerc = .7
iterations = 100
numks = 30
masterAcc = matrix(nrow = iterations, ncol = numks)


for(j in 1:iterations) {
  accs = data.frame(accuracy = numeric(numks), k = numeric(numks))
  trainIndices = sample(1:dim(IPA.Ales)[1],round(splitPerc * dim(IPA.Ales)[1]))
  train = IPA.Ales[trainIndices,]
  test = IPA.Ales[-trainIndices,]
  for(i in 1:numks) {
    classifications = knn(train[,c('IBU','ABV')],test[,c('IBU','ABV')],as.factor(train$type), prob = TRUE, k = i)
    table(as.factor(test$type),classifications)
    CM = confusionMatrix(table(as.factor(test$type),classifications))
    masterAcc[j,i] = CM$overall[1]
  }
}



MeanAcc = colMeans(masterAcc)

plot(seq(1,numks,1),MeanAcc, type = "l")
k <- which.max(MeanAcc)


trainIndices = sample(1:dim(IPA.Ales)[1],round(splitPerc * dim(IPA.Ales)[1]))
train = IPA.Ales[trainIndices,]
test = IPA.Ales[-trainIndices,]

classifications = knn(train[,c('IBU','ABV')],test[,c('IBU','ABV')],as.factor(train$type), prob = TRUE, k = k)
table(as.factor(test$type),classifications)
cfm = confusionMatrix(table(as.factor(test$type),classifications))

Sensitivity = cfm$byClass['Sensitivity']
Specificity = cfm$byClass['Specificity']
Accuracy = cfm$overall['Accuracy']

Accuracy
Sensitivity
Specificity
```
****

## Question 9
### Part 1: Style Predictions from IBU and Ounces (KNN)
**With k = 6 and IBU and Ounce variables, below is the accuracy, sensitivity, and specificity**
```{r, echo=FALSE}
set.seed(100)
#(9)
# Now lets train using KNN but with IBU and ABV used as a seperate variable to classify
mergeData.cleanIBU <- mergeData %>% filter(!is.na(IBU))
typeAles <- cbind(mergeData.cleanIBU, type='Ales', stringsAsFactors=F) %>% filter(grepl('\\bale\\b', Style, ignore.case=T) | grepl('\\bale\\b', beerName, ignore.case=T))

typeIPA <- cbind(mergeData.cleanIBU, type='IPA', stringsAsFactors=F) %>% filter(grepl('\\bIPA\\b', Style, ignore.case=T) | grepl('\\bIPA\\b', beerName, ignore.case=T))

IPA.Ales <- union(typeAles, typeIPA)

iterations = 100

stat = data.frame()

splitPerc = .7 #Training / Test split Percentage

for(j in 1:iterations)
{
  
  trainIndices = sample(1:dim(IPA.Ales)[1],round(splitPerc * dim(IPA.Ales)[1]))
  train = IPA.Ales[trainIndices,]
  test = IPA.Ales[-trainIndices,]
  
  classifications = knn(train[,c('IBU','Ounces')],test[,c('IBU', 'Ounces')],as.factor(train$type), prob = TRUE, k = k)
  table(as.factor(test$type),classifications)
  CM = confusionMatrix(table(as.factor(test$type),classifications))
  
  acc = CM$overall[1]
  spec = CM$byClass[1]
  sens= CM$byClass[2]
  
  stat <- bind_rows(stat,data.frame(c(acc),c(spec),c(sens)))
}

stat <- tibble::rowid_to_column(stat, "Seed")

MeanAcc = mean(stat$c.acc.)
MeanSens = mean(stat$c.sens.)
MeanSpec = mean(stat$c.spec.)

print("Mean Accuracy")
MeanAcc 
print("Mean Sensitivity")
MeanSens
print("Mean Specificity")
MeanSpec
```

### Part 2: Style Predictions from ABV and Ounces (KNN)
**With k = 6 and ABV and Ounce variables, below is the accuracy, sensitivity, and specificity.**
```{r, echo=FALSE}
set.seed(100)
# Do the same for ABV
mergeData.cleanABV <- mergeData %>% filter(!is.na(ABV))
typeAles <- cbind(mergeData.cleanABV, type='Ales', stringsAsFactors=F) %>% filter(grepl('\\bale\\b', Style, ignore.case=T) | grepl('\\bale\\b', beerName, ignore.case=T))

typeIPA <- cbind(mergeData.cleanABV, type='IPA', stringsAsFactors=F) %>% filter(grepl('\\bIPA\\b', Style, ignore.case=T) | grepl('\\bIPA\\b', beerName, ignore.case=T))

IPA.Ales <- union(typeAles, typeIPA)

stat = data.frame()

for(j in 1:iterations)
{
  
  trainIndices = sample(1:dim(IPA.Ales)[1],round(splitPerc * dim(IPA.Ales)[1]))
  train = IPA.Ales[trainIndices,]
  test = IPA.Ales[-trainIndices,]
  
  classifications = knn(train[,c('ABV','Ounces')],test[,c('ABV', 'Ounces')],as.factor(train$type), prob = TRUE, k = k)
  table(as.factor(test$type),classifications)
  CM = confusionMatrix(table(as.factor(test$type),classifications))
  
  acc = CM$overall[1]
  spec = CM$byClass[1]
  sens= CM$byClass[2]
  
  stat <- bind_rows(stat,data.frame(c(acc),c(spec),c(sens)))
}

stat <- tibble::rowid_to_column(stat, "Seed")


MeanAcc = mean(stat$c.acc.)
MeanSens = mean(stat$c.sens.)
MeanSpec = mean(stat$c.spec.)

print("Mean Accuracy")
MeanAcc 
print("Mean Sensitivity")
MeanSens
print("Mean Specificity")
MeanSpec
```

## Conclusion
**We learned that IBU does vary significantly state-to-state, while ABV varies little state-to-state. There is also a positive trend between ABV and IBU of the beers. The higher the ABV the higher the bitterness. Using ABV and IBU, you can somewhat confidently predict the style of beer. The standard size of a beer does not accurately predict the style of beer. We need to ensure future data has more complete IBU fields to get more out of our analysis.**
